---
title: "Working with Missing Data"
output:
  html_document:
    toc: true
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**This tutorial uses fake data simulated using the synthpop package in R.**  

# Setting up R environment
```{r, warning=FALSE, results='hide', message=FALSE}
library(tidyverse) # working with datasets
library(specr) # specification curves
library(mice) # multiple imputation
library(semTools) # missing data information
```

# Loading and prepping data
```{r}
df = read.csv("fake_data.csv") # loading dataset
head(df) # checking data

# correcting variable types (making sure categorical variables are coded as factors)
df = df %>% mutate(
  Gender = as.factor(Gender),
  Race = as.factor(Race),
  Ethnicity =  as.factor(Ethnicity),
  Subject = as.character(Subject)
)
```

# Checking pattern of missing data
```{r}
md.pattern(df, rotate.names = TRUE) # pattern of missing data
mean(is.na(df$Rel_parent)) # checking percentage of missing data
mean(is.na(df$Aggression))
```

# First initial imputations
```{r, warning=FALSE}
ini = mice(df, seed = 329, print = F, m = 5) # dry run of imputations to see how it works

summary(ini) # summary of imputed object
```

# Conducting linear regression on imputed models using MICE
```{r}
fit <- with(ini, lm(Rel_child ~ Trauma)) # linear regression on each imputed dataset separately
summary(pool(fit)) # combining information from multiple datasets
```

# Identifying auxiliary variables
```{r}

## Example for how to check for variables related to missigness. For the sake of today's workshop, we will include all variables in the imputation model (better to err on the side of more!). 

df = df %>% mutate (Parent_missing = is.na(df$Rel_parent)) # create variable which specifies missingness

t.test(df$Age ~ df$Parent_missing) # identifying continuous variables related to missingness

summary(glm(data = df, formula = Parent_missing ~ Gender, family = binomial(link='logit')))  # identifying categorical variables related to missingness
```

# Changing the prediction matrix
```{r}
pred = ini$predictorMatrix # which variables predicted what - can be useful when you don't want certain variables to be predictors. MICE doesn't allow you to remove a variable from being predicted unless you also want to remove it from being a predictor (you will have to create your own prediction matrix for that)
## 1 represents that column variable was used to predict the row variable

pred['Rel_parent', 'Age'] <- 0  # remove specific predictors for specific outcomes
```

# Calculating number of imputed datasets required
```{r}
# percentage of missing data: minimum number of datasets to impute
mean(is.na(df$Rel_parent))

# fraction of missing information (FMI/m should  be less than 0.01)
fmi(df, method = "saturated", group = NULL, ords = NULL,
  varnames = NULL, exclude = NULL, fewImps = FALSE)
```

# Conducting new imputations 
```{r, warning=FALSE}
imp_mypred <- mice(df, predictorMatrix = pred, seed = 329, print=F, maxit = 10, m = 10) # using our own prediction matrix -- change maxit and mice to increase number of iterations and number of imputed data sets respectively

imp_quickpred = mice(df, pred=quickpred(df, mincor=.3), seed = 329, print=F) # the second option is to use pre-set criteria for which variables shouldbe used to predict

plot(imp_mypred) # check convergence of data sets - want them to be free of any trends in later iterations
stripplot(imp_mypred, Aggression~.imp, pch = 20, cex = 2) # check reasonability of imputed values

summary(with(imp_mypred, mean(Rel_parent)))  ## mean from  imputed data set
summary(mean(df$Rel_parent, na.rm = TRUE)) ## mean from non-imputed dataset
```

# Using imputed data set for specification curve analysis
```{r}
# First - save the imputed data so that you don't have to do it again and again
save(imp_mypred, file = "impdata.rda")

# Load the data everytime you want to  use it
load("impdata.rda")

# For specification  curves - we will use the long form of the data
c.long = complete(imp_mypred, 'long') # convert imputed data into long form dataframe
c.broad = complete(imp_mypred, 'broad') # convert imputed data into broad form dataframe
```

## Specification curves on non-imputed data set
```{r}
spec_nonimp = run_specs(df = df,
                    x = c("Trauma"),
                    y = c("Rel_child"),
                    controls = c("Age", "Gender", "Race", "Ethnicity"),
                    model = c("lm"),
                    all.comb = TRUE)
plot_specs(spec_nonimp) # visualize
```

## Specification curves with imputations

### Separate spec curve for each imputed data set
```{r}
  
# running specification curve
spec_imp <- c.long %>% 
  group_by(.imp) %>% # group data based on which imputed data set is
  nest() %>% # creates data frames within data frames
  mutate(specs = purrr::map(data, ~run_specs(df = .,
                    x = c("Trauma"),
                    y = c("Rel_child"),
                    controls = c("Age", "Gender", "Race", "Ethnicity"),
                    model = c("lm"),
                    all.comb = TRUE))) # create a separate specification curve for each imputed data set

```

### Combining results across imputed data sets
```{r}

# unnest specification curves

unnested_specs <- spec_imp %>% unnest(specs) # 'takes out' the specifications from the dataframe

# Function for pooling standard deviation across the imputations

pooled_sd = function(means, sds){
  n = length(means)
  variance_within = (1/n)*sum(sds^2)
  variance_between = var(means)
  variance_total = variance_within + variance_between + (variance_between/n)
  return(sqrt(variance_total))
}

#### grouping and getting mean estimate (rubin's rules)

mean_estimate <- unnested_specs %>% 
  group_by(controls, y, x, model, subsets) %>% 
  summarize(std.error = pooled_sd(means = estimate, 
                                  sds = std.error),
            estimate = mean(estimate)
  ) %>% 
  mutate(conf.low = estimate-2*std.error, 
         conf.high = estimate+2*std.error, 
         statistic = estimate/std.error) %>% 
  ungroup()

plot_specs(mean_estimate) # visualize 
```

### Another example: CBCL ~ Attachment security

```{r}
## running specification curve
spec_imp_attach <- c.long %>% 
  group_by(.imp) %>% # group data based on which imputed data set is
  nest() %>% # creates data frames within data frames
  mutate(specs = purrr::map(data, ~run_specs(df = .,
                    x = c("Rel_child", "Rel_parent"),
                    y = c("Depression"),
                    controls = c("Age", "Gender", "Race", "Ethnicity"),
                    model = c("lm"),
                    all.comb = TRUE))) # create a separate specification curve for each imputed data set

#### unnest specification curves

unnested_specs_attach <- spec_imp_attach %>% unnest(specs) # 'takes out' the specifications from the dataframe

#### grouping and getting mean estimate (rubin's rules)

mean_estimate_attach <- unnested_specs_attach %>% 
  group_by(controls, y, x, model, subsets) %>% 
  summarize(std.error = pooled_sd(means = estimate, 
                                  sds = std.error),
            estimate = mean(estimate)
  ) %>% 
  mutate(conf.low = estimate-2*std.error, 
         conf.high = estimate+2*std.error, 
         statistic = estimate/std.error) %>% 
  ungroup()

plot_specs(mean_estimate_attach) # visualize 
```




